% !TEX root = ../main.tex
\chapter{Repeating Patterns}
\section{Width and Spacing of a Double Slit}

A double slit is placed in the beam path.
The spacing of the two slits is determined by measuring the distance between intensity peaks of the diffraction pattern.
To also determine the width of the individual slits, only a single slit is illuminated and the distance between troughs is measured.
Peaks or troughs are used, as they are more narrow using the respective setup, which makes them easier to measure.
The feature size and error are calculated as shown in \autoref{sec:meas}.

\begin{table}[b!]
	\centering
	\caption{Double Slit Width}
	%\begin{tabular}{SSS}
	\begin{tabular}{
	S[table-format=-1.2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	}
		\toprule
		{$D_\text{screen}$ (\si{\meter})}&	{$S$ (\si{\radian\per\order})}&	{$W$ (\si{\meter})}\\
		\midrule
		2.45&	3.23(13)e-3&	1.96( 8)e-4\\
		2.25&	3.18(15)e-3&	1.99( 9)e-4\\
		2.05&	3.16(16)e-3&	2.00(10)e-4\\
		1.85&	3.20(18)e-3&	1.98(11)e-4\\
		1.65&	3.18(20)e-3&	1.99(12)e-4\\
		\midrule
		{mean}&	3.19( 7)e-3&	1.98( 4)e-4\\
		\bottomrule
	\end{tabular}
\end{table}

\begin{table}[b!]
	\centering
	\caption{Double Slit Spacing}
	\begin{tabular}{
	S[table-format=-1.2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	}
		\toprule
		{$D_\text{screen}$ (\si{\meter})}&	{$S$ (\si{\radian\per\order})}&	{$W$ (\si{\meter})}\\
		\midrule
		2.45&	2.49(13)e-3&	2.54(14)e-4\\
		2.25&	2.51(15)e-3&	2.52(15)e-4\\
		2.05&	2.48(16)e-3&	2.55(16)e-4\\
		1.85&	2.81(18)e-3&	2.25(14)e-4\\
		1.65&	2.49(20)e-3&	2.54(20)e-4\\
		\midrule
		{mean}&	2.55( 7)e-3&	2.49( 7)e-4\\
		\bottomrule
	\end{tabular}
\end{table}

\section{Other Slit Configurations}
\textbf{a)} A second double slit with wider spacing between the slits is used.
The expected result, a more narrow diffraction pattern, is observed at the screen.

\textbf{b)} A triple slit with the same slit width and spacing between the slits is used.
As the feature size remains unchanged, the diffraction pattern should not change in shape.
The third beam adds more chances for destructive interference, so the peaks of the diffraction pattern should become more narrow compared to the double slit.
Both predictions are observed on the screen.
\todo{doesn't sound too good, what do you think?}

\section{Diffraction Grating}
The spacing of the slots in a diffraction grating is analyzed using the same method described earlier.

\begin{table}[b!]
	\centering
	\caption{Diffraction Grating}
	\begin{tabular}{
	S[table-format=-1.2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	S[table-format=-1.3(3)e2, table-align-exponent = false, table-align-uncertainty = false]
	}
		\toprule
		{$D_\text{screen}$ (\si{\meter})}&	{$S$ (\si{\radian\per\order})}&	{$W$ (\si{\meter})}\\
		\midrule
		2.45&	6.00(16)e-3&	1.05( 3)e-4\\
		2.25&	6.08(18)e-3&	1.04( 3)e-4\\
		2.05&	5.90(19)e-3&	1.07( 4)e-4\\
		1.85&	6.00(21)e-3&	1.05( 4)e-4\\
		1.65&	6.08(24)e-3&	1.04( 4)e-4\\
		\midrule
		{mean}&	6.01( 9)e-3&	1.05( 2)e-4\\
		\bottomrule
	\end{tabular}
\end{table}

The factory specification of \num{100} slots per \si{\mm} ($W = \SI{0.1}{\mm}$) lies outside of the calculated error.
\todo{no clue what they mean by 'Welche Rolle spielt die Ausleuchtung?', any ideas?}

\section{Crossed Grating}
A crossed grating is placed in the beam, we probably should have taken a picture of the diffration pattern, but we didn't. So here is a picture of a duck instead.

\begin{figure}[tb]
\centering
\begin{verbatim}
                                              __
                                          ___( o)>
                                          \ <_. )
                                           `---'
\end{verbatim}
\caption{Duck}
\label{fig:duck}
\end{figure}
